Formula translation in C++, Python and modern Fortran: a~case~study~of~the~language~choice tradeoffs

### 2013-07-15 paper/doc/code-cpp-listings.hpp.tex
the new order

\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{//listing00}
\PY{c+c1}{// code licensed under the terms of GNU GPL v3}
\PY{c+c1}{// copyright holder: University of Warsaw}
\PY{c+c1}{//listing01}
\PY{k}{typedef} \PY{k+kt}{double} \PY{n}{real\PYZus{}t}\PY{p}{;}
\PY{c+c1}{//listing02}
\PY{c+cp}{\PYZsh{}}\PY{c+cp}{include \PYZlt{}blitz}\PY{c+cp}{/}\PY{c+cp}{array.h\PYZgt{}}
\PY{k}{using} \PY{n}{arr\PYZus{}t} \PY{o}{=} \PY{n}{blitz}\PY{o}{:}\PY{o}{:}\PY{n}{Array}\PY{o}{\PYZlt{}}\PY{n}{real\PYZus{}t}\PY{p}{,} \PY{l+m+mi}{2}\PY{o}{\PYZgt{}}\PY{p}{;}
\PY{k}{using} \PY{n}{rng\PYZus{}t} \PY{o}{=} \PY{n}{blitz}\PY{o}{:}\PY{o}{:}\PY{n}{Range}\PY{p}{;}
\PY{k}{using} \PY{n}{idx\PYZus{}t} \PY{o}{=} \PY{n}{blitz}\PY{o}{:}\PY{o}{:}\PY{n}{RectDomain}\PY{o}{\PYZlt{}}\PY{l+m+mi}{2}\PY{o}{\PYZgt{}}\PY{p}{;}
\PY{c+c1}{//listing03}
\PY{c+cp}{\PYZsh{}}\PY{c+cp}{define return\PYZus{}macro(expr) \PYZbs{}}
\PY{c+cp}{  -\PYZgt{} decltype(safeToReturn(expr)) \PYZbs{}}
\PY{c+cp}{\PYZob{} return safeToReturn(expr); \PYZcb{} }
\PY{c+c1}{//listing04}
\PY{c+cp}{\PYZsh{}}\PY{c+cp}{include \PYZlt{}boost}\PY{c+cp}{/}\PY{c+cp}{ptr\PYZus{}container}\PY{c+cp}{/}\PY{c+cp}{ptr\PYZus{}vector.hpp\PYZgt{}}
\PY{k}{struct} \PY{n}{arrvec\PYZus{}t} \PY{o}{:} \PY{n}{boost}\PY{o}{:}\PY{o}{:}\PY{n}{ptr\PYZus{}vector}\PY{o}{\PYZlt{}}\PY{n}{arr\PYZus{}t}\PY{o}{\PYZgt{}} 
\PY{p}{\PYZob{}}
  \PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{k}{operator}\PY{p}{[}\PY{p}{]}\PY{p}{(}\PY{k}{const} \PY{k+kt}{int} \PY{n}{i}\PY{p}{)} \PY{k}{const} 
  \PY{p}{\PYZob{}}   
    \PY{k}{return} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{at}\PY{p}{(}\PY{p}{(}\PY{n}{i} \PY{o}{+} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{size}\PY{p}{(}\PY{p}{)}\PY{p}{)} \PY{o}{\PYZpc{}} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{size}\PY{p}{(}\PY{p}{)}\PY{p}{)}\PY{p}{;} 
  \PY{p}{\PYZcb{}}
\PY{p}{\PYZcb{}}\PY{p}{;}
\PY{c+c1}{//listing05}
\PY{k}{struct} \PY{n}{hlf\PYZus{}t} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}} \PY{n}{h}\PY{p}{;}

@@

\PY{k+kr}{inline} \PY{n}{rng\PYZus{}t} \PY{k}{operator}\PY{o}{-}\PY{p}{(}\PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{hlf\PYZus{}t} \PY{o}{\PYZam{}}\PY{p}{)} 
\PY{p}{\PYZob{}} 
  \PY{k}{return} \PY{n}{i}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{;} 
\PY{p}{\PYZcb{}}
\PY{c+c1}{//listing06}
\PY{k}{template}\PY{o}{\PYZlt{}}\PY{k}{class} \PY{n+nc}{n\PYZus{}t}\PY{o}{\PYZgt{}}
\PY{k+kr}{inline} \PY{n}{rng\PYZus{}t} \PY{n}{ext}\PY{p}{(}\PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{r}\PY{p}{,} \PY{k}{const} \PY{n}{n\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{n}\PY{p}{)} \PY{p}{\PYZob{}} 
  \PY{k}{return} \PY{n}{rng\PYZus{}t}\PY{p}{(}
    \PY{p}{(}\PY{n}{r} \PY{o}{-} \PY{n}{n}\PY{p}{)}\PY{p}{.}\PY{n}{first}\PY{p}{(}\PY{p}{)}\PY{p}{,} 
    \PY{p}{(}\PY{n}{r} \PY{o}{+} \PY{n}{n}\PY{p}{)}\PY{p}{.}\PY{n}{last}\PY{p}{(}\PY{p}{)}
  \PY{p}{)}\PY{p}{;} 
\PY{p}{\PYZcb{}} 
\PY{c+c1}{//listing07}
\PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{d}\PY{o}{\PYZgt{}} 
\PY{k+kr}{inline} \PY{n}{idx\PYZus{}t} \PY{n}{pi}\PY{p}{(}\PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}\PY{p}{)}\PY{p}{;}

@@

\PY{k}{template}\PY{o}{\PYZlt{}}\PY{o}{\PYZgt{}}
\PY{k+kr}{inline} \PY{n}{idx\PYZus{}t} \PY{n}{pi}\PY{o}{\PYZlt{}}\PY{l+m+mi}{1}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{)} 
\PY{p}{\PYZob{}}
  \PY{k}{return} \PY{n}{idx\PYZus{}t}\PY{p}{(}\PY{p}{\PYZob{}}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{\PYZcb{}}\PY{p}{)}\PY{p}{;}
\PY{p}{\PYZcb{}}\PY{p}{;} 
\PY{c+c1}{//listing08}
\PY{k}{template}\PY{o}{\PYZlt{}}\PY{k}{class} \PY{n+nc}{bcx\PYZus{}t}\PY{p}{,} \PY{k}{class} \PY{n+nc}{bcy\PYZus{}t}\PY{o}{\PYZgt{}}
\PY{k}{struct} \PY{n}{solver}
\PY{p}{\PYZob{}}
  \PY{c+c1}{// member fields}
  \PY{n}{arrvec\PYZus{}t} \PY{n}{psi}\PY{p}{,} \PY{n}{C}\PY{p}{;}
  \PY{k+kt}{int} \PY{n}{n}\PY{p}{,} \PY{n}{hlo}\PY{p}{;}
  \PY{n}{rng\PYZus{}t} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{;}
  \PY{n}{bcx\PYZus{}t} \PY{n}{bcx}\PY{p}{;}
  \PY{n}{bcy\PYZus{}t} \PY{n}{bcy}\PY{p}{;}

  \PY{c+c1}{// integration logic}
  \PY{k+kt}{void} \PY{n}{solve}\PY{p}{(}\PY{k}{const} \PY{k+kt}{int} \PY{n}{nt}\PY{p}{)} 
  \PY{p}{\PYZob{}}
    \PY{k}{for} \PY{p}{(}\PY{k+kt}{int} \PY{n}{t} \PY{o}{=} \PY{l+m+mi}{0}\PY{p}{;} \PY{n}{t} \PY{o}{\PYZlt{}} \PY{n}{nt}\PY{p}{;} \PY{o}{+}\PY{o}{+}\PY{n}{t}\PY{p}{)} 
    \PY{p}{\PYZob{}}
      \PY{n}{bcx}\PY{p}{.}\PY{n}{fill\PYZus{}halos}\PY{p}{(}\PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{n}{j}\PY{p}{,} \PY{n}{hlo}\PY{p}{)}\PY{p}{)}\PY{p}{;}
      \PY{n}{bcy}\PY{p}{.}\PY{n}{fill\PYZus{}halos}\PY{p}{(}\PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{hlo}\PY{p}{)}\PY{p}{)}\PY{p}{;}
      \PY{n}{advop}\PY{p}{(}\PY{p}{)}\PY{p}{;}
      \PY{n}{cycle}\PY{p}{(}\PY{p}{)}\PY{p}{;}
    \PY{p}{\PYZcb{}}
  \PY{p}{\PYZcb{}}
\PY{p}{\PYZcb{}}\PY{p}{;}
\PY{c+c1}{//listing09}
\PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{d}\PY{o}{\PYZgt{}}
\PY{k}{struct} \PY{n}{cyclic}
\PY{p}{\PYZob{}}
  \PY{c+c1}{// member fields}
  \PY{n}{rng\PYZus{}t} \PY{n}{left\PYZus{}halo}\PY{p}{,} \PY{n}{rght\PYZus{}halo}\PY{p}{;}
  \PY{n}{rng\PYZus{}t} \PY{n}{left\PYZus{}edge}\PY{p}{,} \PY{n}{rght\PYZus{}edge}\PY{p}{;}\PY{p}{;}

  \PY{c+c1}{// method invoked by the solver}
  \PY{k+kt}{void} \PY{n}{fill\PYZus{}halos}\PY{p}{(}\PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{a}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}\PY{p}{)}

@@

    \PY{n}{a}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{left\PYZus{}halo}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)} \PY{o}{=} \PY{n}{a}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{rght\PYZus{}edge}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{;}     
    \PY{n}{a}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{rght\PYZus{}halo}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)} \PY{o}{=} \PY{n}{a}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{left\PYZus{}edge}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{;}     

@@

@@

@@

@@

\PY{k}{struct} \PY{n}{solver\PYZus{}donorcell} \PY{o}{:} \PY{n}{solver}\PY{o}{\PYZlt{}}\PY{n}{bcx\PYZus{}t}\PY{p}{,} \PY{n}{bcy\PYZus{}t}\PY{o}{\PYZgt{}} 
  \PY{n}{solver\PYZus{}donorcell}\PY{p}{(}\PY{k+kt}{int} \PY{n}{nx}\PY{p}{,} \PY{k+kt}{int} \PY{n}{ny}\PY{p}{)} \PY{o}{:}
    \PY{n}{solver}\PY{o}{\PYZlt{}}\PY{n}{bcx\PYZus{}t}\PY{p}{,} \PY{n}{bcy\PYZus{}t}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{nx}\PY{p}{,} \PY{n}{ny}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}
  \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}

  \PY{k+kt}{void} \PY{n}{advop}\PY{p}{(}\PY{p}{)}
    \PY{n}{donorcell\PYZus{}op}\PY{p}{(}
      \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{psi}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{n}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{C}\PY{p}{,} 
      \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{j}

@@

\PY{c+c1}{//listing17}

@@

\PY{c+c1}{//listing18}

@@

\PY{c+c1}{//listing19}
\PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{n\PYZus{}iters}\PY{p}{,} \PY{k}{class} \PY{n+nc}{bcx\PYZus{}t}\PY{p}{,} \PY{k}{class} \PY{n+nc}{bcy\PYZus{}t}\PY{o}{\PYZgt{}}
\PY{k}{struct} \PY{n}{solver\PYZus{}mpdata} \PY{o}{:} \PY{n}{solver}\PY{o}{\PYZlt{}}\PY{n}{bcx\PYZus{}t}\PY{p}{,} \PY{n}{bcy\PYZus{}t}\PY{o}{\PYZgt{}}

@@

  \PY{n}{arrvec\PYZus{}t} \PY{n}{tmp}\PY{p}{[}\PY{l+m+mi}{2}\PY{p}{]}\PY{p}{;}
  \PY{n}{rng\PYZus{}t} \PY{n}{im}\PY{p}{,} \PY{n}{jm}\PY{p}{;}

  \PY{c+c1}{// ctor}
  \PY{n}{solver\PYZus{}mpdata}\PY{p}{(}\PY{k+kt}{int} \PY{n}{nx}\PY{p}{,} \PY{k+kt}{int} \PY{n}{ny}\PY{p}{)} \PY{o}{:} 
    \PY{n}{solver}\PY{o}{\PYZlt{}}\PY{n}{bcx\PYZus{}t}\PY{p}{,} \PY{n}{bcy\PYZus{}t}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{nx}\PY{p}{,} \PY{n}{ny}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}\PY{p}{,} 
    \PY{n}{im}\PY{p}{(}\PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{.}\PY{n}{first}\PY{p}{(}\PY{p}{)} \PY{o}{-} \PY{l+m+mi}{1}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{i}\PY{p}{.}\PY{n}{last}\PY{p}{(}\PY{p}{)}\PY{p}{)}\PY{p}{,}
    \PY{n}{jm}\PY{p}{(}\PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{j}\PY{p}{.}\PY{n}{first}\PY{p}{(}\PY{p}{)} \PY{o}{-} \PY{l+m+mi}{1}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{j}\PY{p}{.}\PY{n}{last}\PY{p}{(}\PY{p}{)}\PY{p}{)}
  \PY{p}{\PYZob{}}
    \PY{k+kt}{int} \PY{n}{n\PYZus{}tmp} \PY{o}{=} \PY{n}{n\PYZus{}iters} \PY{o}{\PYZgt{}} \PY{l+m+mi}{2} \PY{o}{?} \PY{l+m+mi}{2} \PY{o}{:} \PY{l+m+mi}{1}\PY{p}{;}
    \PY{k}{for} \PY{p}{(}\PY{k+kt}{int} \PY{n}{n} \PY{o}{=} \PY{l+m+mi}{0}\PY{p}{;} \PY{n}{n} \PY{o}{\PYZlt{}} \PY{n}{n\PYZus{}tmp}\PY{p}{;} \PY{o}{+}\PY{o}{+}\PY{n}{n}\PY{p}{)}
    \PY{p}{\PYZob{}}
      \PY{n}{tmp}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{.}\PY{n}{push\PYZus{}back}\PY{p}{(}\PY{k}{new} \PY{n}{arr\PYZus{}t}\PY{p}{(}
        \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{C}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{.}\PY{n}{domain}\PY{p}{(}\PY{p}{)}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{C}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{.}\PY{n}{domain}\PY{p}{(}\PY{p}{)}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{)}
      \PY{p}{)}\PY{p}{;}
      \PY{n}{tmp}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{.}\PY{n}{push\PYZus{}back}\PY{p}{(}\PY{k}{new} \PY{n}{arr\PYZus{}t}\PY{p}{(}
        \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{C}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{.}\PY{n}{domain}\PY{p}{(}\PY{p}{)}\PY{p}{[}\PY{l+m+mi}{0}\PY{p}{]}\PY{p}{,} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{C}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{.}\PY{n}{domain}\PY{p}{(}\PY{p}{)}\PY{p}{[}\PY{l+m+mi}{1}\PY{p}{]}\PY{p}{)}
      \PY{p}{)}\PY{p}{;}
    \PY{p}{\PYZcb{}}
  \PY{p}{\PYZcb{}}

>>>>

\begin{Verbatim}[commandchars=\\\{\}]
\PY{c+c1}{//listing00}
\PY{c+c1}{// code licensed under the terms of GNU GPL v3}
\PY{c+c1}{// copyright holder: University of Warsaw}
\PY{c+c1}{//listing01}
\PY{k}{using} \PY{n}{real\PYZus{}t} \PY{o}{=} \PY{k+kt}{double}\PY{p}{;}
\PY{c+c1}{//listing02}
\PY{c+cp}{\PYZsh{}}\PY{c+cp}{include \PYZlt{}blitz}\PY{c+cp}{/}\PY{c+cp}{array.h\PYZgt{}}
\PY{k}{using} \PY{n}{arr\PYZus{}t} \PY{o}{=} \PY{n}{blitz}\PY{o}{:}\PY{o}{:}\PY{n}{Array}\PY{o}{\PYZlt{}}\PY{n}{real\PYZus{}t}\PY{p}{,} \PY{l+m+mi}{2}\PY{o}{\PYZgt{}}\PY{p}{;}
\PY{k}{using} \PY{n}{rng\PYZus{}t} \PY{o}{=} \PY{n}{blitz}\PY{o}{:}\PY{o}{:}\PY{n}{Range}\PY{p}{;}
\PY{k}{using} \PY{n}{idx\PYZus{}t} \PY{o}{=} \PY{n}{blitz}\PY{o}{:}\PY{o}{:}\PY{n}{RectDomain}\PY{o}{\PYZlt{}}\PY{l+m+mi}{2}\PY{o}{\PYZgt{}}\PY{p}{;}
\PY{c+c1}{//listing03}
\PY{c+cp}{\PYZsh{}}\PY{c+cp}{define return\PYZus{}macro(expr) \PYZbs{}}
\PY{c+cp}{  -\PYZgt{} decltype(safeToReturn(expr)) \PYZbs{}}
\PY{c+cp}{\PYZob{} return safeToReturn(expr); \PYZcb{} }
\PY{c+c1}{//listing04}
\PY{c+cp}{\PYZsh{}}\PY{c+cp}{include \PYZlt{}boost}\PY{c+cp}{/}\PY{c+cp}{ptr\PYZus{}container}\PY{c+cp}{/}\PY{c+cp}{ptr\PYZus{}vector.hpp\PYZgt{}}
\PY{k}{struct} \PY{n}{arrvec\PYZus{}t} \PY{o}{:} \PY{n}{boost}\PY{o}{:}\PY{o}{:}\PY{n}{ptr\PYZus{}vector}\PY{o}{\PYZlt{}}\PY{n}{arr\PYZus{}t}\PY{o}{\PYZgt{}} 
\PY{p}{\PYZob{}}
  \PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{k}{operator}\PY{p}{[}\PY{p}{]}\PY{p}{(}\PY{k}{const} \PY{k+kt}{int} \PY{n}{i}\PY{p}{)} \PY{k}{const} 
  \PY{p}{\PYZob{}}   
    \PY{k}{return} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{at}\PY{p}{(}\PY{p}{(}\PY{n}{i} \PY{o}{+} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{size}\PY{p}{(}\PY{p}{)}\PY{p}{)} \PY{o}{\PYZpc{}} \PY{k}{this}\PY{o}{-}\PY{o}{\PYZgt{}}\PY{n}{size}\PY{p}{(}\PY{p}{)}\PY{p}{)}\PY{p}{;} 
  \PY{p}{\PYZcb{}}
\PY{p}{\PYZcb{}}\PY{p}{;}
\PY{c+c1}{//listing05}
\PY{k}{struct} \PY{n}{hlf\PYZus{}t} \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}} \PY{n}{h}\PY{p}{;}

@@

\PY{k+kr}{inline} \PY{n}{rng\PYZus{}t} \PY{k}{operator}\PY{o}{-}\PY{p}{(}\PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{hlf\PYZus{}t} \PY{o}{\PYZam{}}\PY{p}{)} 
\PY{p}{\PYZob{}} 
  \PY{k}{return} \PY{n}{i}\PY{o}{-}\PY{l+m+mi}{1}\PY{p}{;} 
\PY{p}{\PYZcb{}}
\PY{c+c1}{//listing06}
\PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{d}\PY{o}{\PYZgt{}} 
\PY{k+kr}{inline} \PY{n}{idx\PYZus{}t} \PY{n}{pi}\PY{p}{(}\PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}\PY{p}{)}\PY{p}{;}

@@

\PY{k}{template}\PY{o}{\PYZlt{}}\PY{o}{\PYZgt{}}
\PY{k+kr}{inline} \PY{n}{idx\PYZus{}t} \PY{n}{pi}\PY{o}{\PYZlt{}}\PY{l+m+mi}{1}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{i}\PY{p}{)} 

@@

  \PY{k}{return} \PY{n}{idx\PYZus{}t}\PY{p}{(}\PY{p}{\PYZob{}}\PY{n}{i}\PY{p}{,}\PY{n}{j}\PY{p}{\PYZcb{}}\PY{p}{)}\PY{p}{;}

@@

\PY{c+c1}{//listing07}

@@

\PY{c+c1}{//listing08}

@@

\PY{c+c1}{//listing09}

@@

@@

@@

@@

\PY{k}{template}\PY{o}{\PYZlt{}}\PY{k}{class} \PY{n+nc}{n\PYZus{}t}\PY{o}{\PYZgt{}}
\PY{k+kr}{inline} \PY{n}{rng\PYZus{}t} \PY{n}{ext}\PY{p}{(}\PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{r}\PY{p}{,} \PY{k}{const} \PY{n}{n\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{n}\PY{p}{)} \PY{p}{\PYZob{}} 
  \PY{k}{return} \PY{n}{rng\PYZus{}t}\PY{p}{(}
    \PY{p}{(}\PY{n}{r} \PY{o}{-} \PY{n}{n}\PY{p}{)}\PY{p}{.}\PY{n}{first}\PY{p}{(}\PY{p}{)}\PY{p}{,} 
    \PY{p}{(}\PY{n}{r} \PY{o}{+} \PY{n}{n}\PY{p}{)}\PY{p}{.}\PY{n}{last}\PY{p}{(}\PY{p}{)}
\PY{k}{struct} \PY{n}{solver}

@@

  \PY{n}{arrvec\PYZus{}t} \PY{n}{psi}\PY{p}{,} \PY{n}{C}\PY{p}{;}
  \PY{k+kt}{int} \PY{n}{n}\PY{p}{,} \PY{n}{hlo}\PY{p}{;}
  \PY{n}{rng\PYZus{}t} \PY{n}{i}\PY{p}{,} \PY{n}{j}\PY{p}{;}
  \PY{n}{bcx\PYZus{}t} \PY{n}{bcx}\PY{p}{;}
  \PY{n}{bcy\PYZus{}t} \PY{n}{bcy}\PY{p}{;}

  \PY{c+c1}{// integration logic}
  \PY{k+kt}{void} \PY{n}{solve}\PY{p}{(}\PY{k}{const} \PY{k+kt}{int} \PY{n}{nt}\PY{p}{)} 
  \PY{p}{\PYZob{}}
    \PY{k}{for} \PY{p}{(}\PY{k+kt}{int} \PY{n}{t} \PY{o}{=} \PY{l+m+mi}{0}\PY{p}{;} \PY{n}{t} \PY{o}{\PYZlt{}} \PY{n}{nt}\PY{p}{;} \PY{o}{+}\PY{o}{+}\PY{n}{t}\PY{p}{)} 
    \PY{p}{\PYZob{}}
      \PY{n}{bcx}\PY{p}{.}\PY{n}{fill\PYZus{}halos}\PY{p}{(}\PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{n}{j}\PY{p}{,} \PY{n}{hlo}\PY{p}{)}\PY{p}{)}\PY{p}{;}
      \PY{n}{bcy}\PY{p}{.}\PY{n}{fill\PYZus{}halos}\PY{p}{(}\PY{n}{psi}\PY{p}{[}\PY{n}{n}\PY{p}{]}\PY{p}{,} \PY{n}{ext}\PY{p}{(}\PY{n}{i}\PY{p}{,} \PY{n}{hlo}\PY{p}{)}\PY{p}{)}\PY{p}{;}
      \PY{n}{advop}\PY{p}{(}\PY{p}{)}\PY{p}{;}
      \PY{n}{cycle}\PY{p}{(}\PY{p}{)}\PY{p}{;}
    \PY{p}{\PYZcb{}}
  \PY{p}{\PYZcb{}}
\PY{p}{\PYZcb{}}\PY{p}{;}
\PY{c+c1}{//listing17}
\PY{k}{template}\PY{o}{\PYZlt{}}\PY{k+kt}{int} \PY{n}{d}\PY{o}{\PYZgt{}}
\PY{k}{struct} \PY{n}{cyclic}
\PY{p}{\PYZob{}}
  \PY{c+c1}{// member fields}
  \PY{n}{rng\PYZus{}t} \PY{n}{left\PYZus{}halo}\PY{p}{,} \PY{n}{rght\PYZus{}halo}\PY{p}{;}
  \PY{n}{rng\PYZus{}t} \PY{n}{left\PYZus{}edge}\PY{p}{,} \PY{n}{rght\PYZus{}edge}\PY{p}{;}\PY{p}{;}

  \PY{c+c1}{// method invoked by the solver}
  \PY{k+kt}{void} \PY{n}{fill\PYZus{}halos}\PY{p}{(}\PY{k}{const} \PY{n}{arr\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{a}\PY{p}{,} \PY{k}{const} \PY{n}{rng\PYZus{}t} \PY{o}{\PYZam{}}\PY{n}{j}\PY{p}{)}
  \PY{p}{\PYZob{}}
    \PY{n}{a}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{left\PYZus{}halo}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)} \PY{o}{=} \PY{n}{a}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{rght\PYZus{}edge}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{;}     
    \PY{n}{a}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{rght\PYZus{}halo}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)} \PY{o}{=} \PY{n}{a}\PY{p}{(}\PY{n}{pi}\PY{o}{\PYZlt{}}\PY{n}{d}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{left\PYZus{}edge}\PY{p}{,} \PY{n}{j}\PY{p}{)}\PY{p}{)}\PY{p}{;}     
  \PY{p}{\PYZcb{}}
\PY{p}{\PYZcb{}}\PY{p}{;}
\PY{c+c1}{//listing18}
\PY{k}{template}\PY{o}{\PYZlt{}}\PY{k}{class} \PY{n+nc}{bcx\PYZus{}t}\PY{p}{,} \PY{k}{class} \PY{n+nc}{bcy\PYZus{}t}\PY{o}{\PYZgt{}}
\PY{k}{struct} \PY{n}{solver\PYZus{}donorcell} \PY{o}{:} \PY{n}{solver}\PY{o}{\PYZlt{}}\PY{n}{bcx\PYZus{}t}\PY{p}{,} \PY{n}{bcy\PYZus{}t}\PY{o}{\PYZgt{}} 
\PY{p}{\PYZob{}}
  \PY{n}{solver\PYZus{}donorcell}\PY{p}{(}\PY{k+kt}{int} \PY{n}{nx}\PY{p}{,} \PY{k+kt}{int} \PY{n}{ny}\PY{p}{)} \PY{o}{:}
    \PY{n}{solver}\PY{o}{\PYZlt{}}\PY{n}{bcx\PYZus{}t}\PY{p}{,} \PY{n}{bcy\PYZus{}t}\PY{o}{\PYZgt{}}\PY{p}{(}\PY{n}{nx}\PY{p}{,} \PY{n}{ny}\PY{p}{,} \PY{l+m+mi}{1}\PY{p}{)}
  \PY{p}{\PYZob{}}\PY{p}{\PYZcb{}}
